---
title: "Savannah Work"
author: "Savannah Flanagan"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
library(gtsummary)
library(emulator)
```
# Preprocessing and EDA

## Initial Preprocessing of Data
- Filled in some missing is_first_child values
- Converted all variables to factors
- Dropped NAs

```{r}
set.seed(4)
test_df=read_csv('Project_1_data.csv') %>% janitor::clean_names() 

test_df=test_df %>%  
  mutate(is_first_child=case_when(
    is.na(is_first_child) &nr_siblings==0 ~ 'yes',
    TRUE ~is_first_child
  )) %>% 
        mutate(gender=as.factor(gender),
               ethnic_group=as.factor(ethnic_group),
               parent_educ=as.factor(parent_educ),
               lunch_type=as.factor(lunch_type),
               test_prep=as.factor(test_prep),
               parent_marital_status=as.factor(parent_marital_status),
               practice_sport=as.factor(practice_sport),
               is_first_child=as.factor(is_first_child),
               transport_means=as.factor(transport_means),
               wkly_study_hours=as.factor(wkly_study_hours)) %>% drop_na()
 
```
## EDA
- Look at the general summary stats of all variables
- Look at the distributions of all outcomes (test scores)
```{r}
summary(test_df)

test_df %>% 
  ggplot(aes(x=math_score))+geom_histogram()
test_df %>% 
  ggplot(aes(x=reading_score))+geom_histogram()
test_df %>% 
  ggplot(aes(x=writing_score))+geom_histogram()
```
## Testing for Variable Associations
- Look for collinearity
- Chi-squared test results show a significant relationship between:
     - gender and nr_siblings
     - is_first_child and nr_siblings
- Because of these relationships, we decided to drop nr_siblings to reduce collinearity

```{r}

# Loop through pairs of columns and run the chi-square test

for (i in 1:(11 - 1)) {
  for (j in (i + 1):11) {
    # Get the two columns to test
    col1 <- test_df[[i]]
    col2 <- test_df[[j]]
    
    contingency_table <- table(col1, col2)
    
    # Perform the chi-square test
    chi_test <- chisq.test(contingency_table)
    
    #PRint the results 
    if(chi_test$p.value<.05){
      
      
      cat("Chi-square test for columns", colnames(test_df)[i], "and", colnames(test_df)[j], "\n")
    print(chi_test)
    

    }
    
  }
}

ggplot(test_df, aes(x = nr_siblings, fill = gender)) +
      geom_bar(position = "fill") +
      labs(y = "Proportion", title = "Pairwise Relationship")
ggplot(test_df, aes(x = nr_siblings, fill = is_first_child)) +
      geom_bar(position = "fill") +
      labs(y = "Proportion", title = "Pairwise Relationship")

```
## Explore Models for Predicting Scores Based on the Main Covariates

### Writing Score Model

Make df and fit model
```{r}
set.seed(4)
df_4_mdl_write=test_df %>% dplyr::select(gender,ethnic_group,parent_educ,lunch_type,test_prep,parent_marital_status,practice_sport,is_first_child,transport_means,wkly_study_hours,writing_score)
full_fit_write=lm(writing_score ~ .^2,data=df_4_mdl_write) # includes interaction effects
```

Use stepwise regression to simplify the model
```{r}
set.seed(4)
stepwise_model_write <- step(full_fit_write, direction = "both", trace=0)

summary(stepwise_model_write)
```
Removing insignificant interaction terms from the writing model:

```{r}
set.seed(4)
full_fit_write_minus_insig=lm(writing_score ~ gender + ethnic_group + parent_educ +  lunch_type + test_prep + parent_marital_status + practice_sport + is_first_child + transport_means + wkly_study_hours,data=df_4_mdl_write) # has no significant interaction effects

summary(full_fit_write_minus_insig)
plot(full_fit_write_minus_insig)
```


### Math Score Model

Make df and fit model
```{r}
set.seed(4)
df_4_mdl_math=test_df %>% dplyr::select(gender,ethnic_group,parent_educ,lunch_type,test_prep,parent_marital_status,practice_sport,is_first_child,transport_means,wkly_study_hours,math_score)
full_fit_math=lm(math_score ~ .^2,data=df_4_mdl_math)
```


Use stepwise regression to simplify the model
```{r}
set.seed(4)
stepwise_model_math <- step(full_fit_math, direction = "both", trace = 0)

summary(stepwise_model_math)
```
Removing insignificant interaction terms from the math model:

```{r}
set.seed(4)
full_fit_math_minus_insig=lm(math_score ~ gender + ethnic_group + parent_educ +  lunch_type + test_prep + parent_marital_status + practice_sport + is_first_child + transport_means + wkly_study_hours + lunch_type*practice_sport,data=df_4_mdl_math) # has no significant interaction effects

summary(full_fit_math_minus_insig)
plot(full_fit_math_minus_insig)
```



### Reading Score 

Make df and fit model
```{r}
set.seed(4)
df_4_mdl_read=test_df %>% dplyr::select(gender,ethnic_group,parent_educ,lunch_type,test_prep,parent_marital_status,practice_sport,is_first_child,transport_means,wkly_study_hours,reading_score)
full_fit_read=lm(reading_score ~ .^2,data=df_4_mdl_read)
```

Use stepwise regression to simplify the model
```{r}
set.seed(4)
stepwise_model_read <- step(full_fit_read, direction = "both", trace = 0)

summary(stepwise_model_read)
```

```{r}
set.seed(4)
full_fit_read_minus_insig=lm(reading_score ~ gender + ethnic_group + parent_educ +  lunch_type + test_prep + parent_marital_status + practice_sport + is_first_child + transport_means + wkly_study_hours,data=df_4_mdl_read) # has no significant interaction effects

summary(full_fit_read_minus_insig)
plot(full_fit_read_minus_insig)
```


## Explore Diagnostics for These Models (with Insignificent Interaction Terms)

### Writing Score Model Diagnostics

```{r}
plot(stepwise_model_write)
```
### Math Score Model Diagnostics

```{r}
plot(stepwise_model_math)
```
### Reading Score Model Diagnostics

```{r}
plot(stepwise_model_read)
```


## Explore Models for Predicting Scores Based on the Main Covariates and Other Scores

### Writing Score Model with Other Scores
```{r}
set.seed(4)
df_4_mdl_write2=test_df %>% dplyr::select(gender,ethnic_group,parent_educ,lunch_type,test_prep,parent_marital_status,practice_sport,is_first_child,transport_means,wkly_study_hours,writing_score,math_score,reading_score)
full_fit_write2=lm(writing_score ~ .^2,data=df_4_mdl_write2)
```


```{r}
set.seed(4)
stepwise_model_write2 <- step(full_fit_write2, direction = "both", trace = 0)
summary(stepwise_model_write2)
```
Diagnostics
```{r}
plot(stepwise_model_write2)
```

### Math Score Model with Other Scores

```{r}
set.seed(4)
df_4_mdl_math2=test_df %>% dplyr::select(gender,ethnic_group,parent_educ,lunch_type,test_prep,parent_marital_status,practice_sport,is_first_child,transport_means,wkly_study_hours,math_score,writing_score,reading_score)
full_fit_math2=lm(math_score ~ .^2,data=df_4_mdl_math2)
```

```{r}
set.seed(4)
stepwise_model_math2 <- step(full_fit_math2, direction = "both", trace = 0)
summary(stepwise_model_math2)
```
Diagnostics
```{r}
plot(stepwise_model_math2)
```


### Reading Score Score Model with Other Scores

```{r}
set.seed(4)
df_4_mdl_read2=test_df %>% dplyr::select(gender,ethnic_group,parent_educ,lunch_type,test_prep,parent_marital_status,practice_sport,is_first_child,transport_means,wkly_study_hours,reading_score,writing_score,math_score)
full_fit_read2=lm(reading_score ~ .^2,data=df_4_mdl_read2)
```

```{r}
set.seed(4)
stepwise_model_read2 <- step(full_fit_read2, direction = "both", trace = 0)

summary(stepwise_model_read2)
```

Diagnostics
```{r}
plot(stepwise_model_read2)
```


## Removing the insignificant terms from the model that predicts with covariates and the other scores

Writing
```{r}
set.seed(4)

writing_without_insig_interactions <- lm(writing_score ~ gender + ethnic_group + parent_educ +     lunch_type + test_prep + parent_marital_status + practice_sport +    is_first_child + transport_means + wkly_study_hours + reading_score + math_score + transport_means:reading_score + math_score:reading_score, data=df_4_mdl_read2)

summary(writing_without_insig_interactions)
plot(writing_without_insig_interactions)
```

Math
```{r}
set.seed(4)

math_without_insig_interactions <- lm(math_score ~ gender + ethnic_group + parent_educ +     lunch_type + test_prep + parent_marital_status + practice_sport +    is_first_child + transport_means + wkly_study_hours + reading_score + writing_score + lunch_type:is_first_child + test_prep:transport_means + wkly_study_hours:reading_score, data=df_4_mdl_read2)

summary(math_without_insig_interactions)
plot(math_without_insig_interactions)
```

Reading
```{r}
set.seed(4)

reading_without_insig_interactions <- lm(reading_score ~ gender + ethnic_group + parent_educ +     lunch_type + test_prep + parent_marital_status + practice_sport +    is_first_child + transport_means + wkly_study_hours + writing_score + math_score + is_first_child:transport_means + transport_means:writing_score, data=df_4_mdl_read2)

summary(reading_without_insig_interactions)
plot(reading_without_insig_interactions)
```





XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
I removed the interaction terms that weren't present in both models (covariates only and with other score) as well as weren't another variable interacting with the other scores
```{r}
set.seed(4)

reading_without_interactions <- lm(reading_score ~ gender + ethnic_group + parent_educ +     lunch_type + test_prep + parent_marital_status + practice_sport +    is_first_child + transport_means + wkly_study_hours + writing_score + math_score + gender:writing_score + gender:math_score + ethnic_group:transport_means +     ethnic_group:writing_score + parent_educ:math_score +   lunch_type:wkly_study_hours + 
parent_marital_status:is_first_child +     parent_marital_status:wkly_study_hours + parent_marital_status:writing_score + practice_sport:writing_score + practice_sport:reading_score + transport_means:writing_score + wkly_study_hours:writing_score + wkly_study_hours:math_score, data=df_4_mdl_read2)

summary(reading_without_interactions)
```

Now, take the model above and remove insignficant interaction terms (if some of the categories of a variable are insignificant, remove the whole interaction term)
```{r}
set.seed(4)

reading_without_interactions_or_insig <- lm(reading_score ~ gender + ethnic_group + parent_educ +     lunch_type + test_prep + parent_marital_status + practice_sport +    is_first_child + transport_means + wkly_study_hours + reading_score + math_score + gender:writing_score +  practice_sport:writing_score + practice_sport:reading_score, data=df_4_mdl_read2)

summary(reading_without_interactions_or_insig)
```
Take that last model, and just look at diagnostic plots
```{r}
plot(reading_without_interactions)
```